Unveiling Exception Handling Bug Hazards in Android based on GitHub and Google Code Issues

### 2015-02-15 android-stacks.tex
Sharpened conclusions on exception types (IV.B)

In recent years, we have witnessed an astonishing increase in the number of
mobile applications. These applications extend phones capabilities 
far beyond basic calls and textual messages. Such applications, however,
have to face an increasing number of threats to application robustness
 arising from failures in the underlying middleware and hardware (e.g., camera, sensors);
failures in third party services and libraries; compatibility issues~\cite{McDon13}; 
 memory and battery restrictions; and noisy external resources~\cite{Zhang12} 
(e.g., wireless connections,GPS, bluetooth).

@@

 A common way of  propagating exceptions in Java programs is through exception wrapping
 (also known as exception chaining): One exception 
is caught and wrapped in another which is then thrown instead. Figure~\ref{fig:wrapping} shows 
an exception stack trace which illustrates such exception wrapping. 
For simplicity, in this paper we will refer to ``exception stack trace'' as as just stack trace.
The bottom part of the stack trace is the \emph{root exception}, which indicates
the first reason (root cause) for the exception thrown (in this case, the computer run out of
memory). The top part of the stack trace indicates the location of the exception
manifestation (which we will refer to as the \emph{exception wrapper} in this paper). The
execution flow  between the root exception and the wrapper may
include other intermediate exception wrappers. In all levels, the exception
\emph{signaler}, is the method that threw the exception, represented on the
stack trace as the first method call below the exception declaration.

@@

\emph{\textbf{Inspecting Exception Interfaces.}}  According to the best practices mentioned before, 
 \emph{explicitly thrown} runtime exceptions 
should be documented as part of the \emph{exception interface} of libraries/framework reusable
methods. To investigate the conformance to this practice, we firstly filtered out all the exceptions implicitly
 signaled by  the runtime environment (due to programming mistakes) - since these exceptions 
should not be documented on the method signature.  Then we inspected
the code for each method 
(defined either on Android  Aplication Framework or on third-party libraries) 
explicitly signaling a runtime exception. Table~\ref{tab:runtimeinterface} presents the results of this inspection. 
We could find 79 methods (both from libraries and Android platform) that  explicitly threw a runtime exception 
without listing such exception on the \emph{exception interface} (i.e., using 
throws clause on method signature). From this set only one method (defined on a library)
included an @throws tag on Javadoc - reporting that the given runtime exception
could be thrown in some contidions. These methods were responsible for 
267 exception stack traces mined in this study.

This result is aligned with the results of other two studies~\cite{sacramento2006unchecked,kechagia2014}. 
One conducted by  Sacramento et al~\cite{sacramento2006unchecked} - which observed that the
runtime exceptions in .Net programs are most often not documented. And other conducted by 
Kechagia and Spinellis~\cite{kechagia2014} which discovered a set of methods 
on the Android API which do not documented runtime exceptions. One limitation 
of the work of Kechagia and Spinellis is that it did not filter out exceptions that,  
although runtime, should not be documented because they were implicitly signaled by the 
JVM due to resource restrictions or violations on Java semantic constraints.
When explicitly signaling a runtime exception and not documenting it, 
the developer imposes a threat to system robustness, specially
when such exceptions are thrown by a third party code (e.g. libraries, or framework utility code)
 invoked inside the application. In such cases the developer usually does not have access to 
the source code. Hence in the absence of the exception documentation it is very difficult or even impossible
 for the client to design the application to deal with ``unforeseen" runtime exceptions. As a consequence, the
 undocumented runtime exception may remain uncaught and lead to system crashes.

\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
The study revealed: 267 exception stack traces were caused by ``undocumented runtime exceptions" 
thrown by libraries or the Android platform.
}
\end{minipage}}

@@

 %\begin{tabular}{|p{7.9cm} p{0.1cm}|}
\begin{tabular}{lrrrr}
    \hline
 \bfseries{Origin}   &  \bfseries{stacks}  & \bfseries{signaler methods} &  \bfseries{throws clause}  &  \bfseries{\@throws}  \\ 
\hline					
Libraries	& 	205	 & 	44	& 0	& 1	\\
Android  	&	62 &	35	&0 &  0	\\					
\hline					
\bfseries{All}	&	267 &	79 &  0  & 1\\
\hline					
  \end{tabular}
\caption{Absence of exception interfaces on methods.}
\label{tab:runtimeinterface}
\end{table}

>>>>

In recent years, we have witnessed an astonishing increase in the number of
mobile applications. These applications extend phones capabilities 
far beyond basic calls and textual messages. They must, however,
face an increasing number of threats to application robustness
 arising from failures in the underlying middleware and hardware (e.g., camera, sensors);
failures in third party services and libraries; compatibility issues~\cite{McDon13}; 
 memory and battery restrictions; and noisy external resources~\cite{Zhang12} 
(e.g., wireless connections,GPS, bluetooth).

@@

 A common way of  propagating exceptions in Java programs is through
 exception wrapping (also called chaining):
One exception 
is caught and wrapped in another which is then thrown instead. Figure~\ref{fig:wrapping} shows 
an exception stack trace which illustrates such exception wrapping. 
For simplicity, in this paper we will refer to ``exception stack trace'' as as just stack trace.
The bottom part of the stack trace is the \emph{root exception}, which indicates
the first reason (root cause) for the exception thrown (in this case, the computer run out of
memory). The top part of the stack trace indicates the location of the exception
manifestation (which we will refer to as the \emph{exception wrapper} in this paper). The
execution flow  between the root exception and the wrapper may
include other intermediate exception wrappers. In all levels, the exception
\emph{signaler}, is the method that threw the exception, represented on the
stack trace as the first method call below the exception declaration.

@@

\finding{Almost two thirds of all crashes come from run time
  exceptions. Most of these originate from the application layer.}

\emph{\textbf{Inspecting Exception Interfaces.}}  According to the best practices mentioned before, 
 \emph{explicitly thrown} runtime exceptions 
should be documented as part of the \emph{exception interface} of libraries/framework reusable
methods. To investigate the conformance to this practice, we firstly filtered out all the exceptions implicitly
 signaled by  the runtime environment (due to programming mistakes) - since these exceptions 
should not be documented on the method signature.  Then we inspected
the code for each method 
(defined either in the Android  Aplication Framework or in third-party libraries) 
explicitly signaling a runtime exception.

Table~\ref{tab:runtimeinterface} presents the results of this inspection. 
We found 79 methods (both from libraries and the Android platform) that  explicitly threw a runtime exception 
without listing it on the \emph{exception interface} (i.e., using 
throws clause in the method signature). From this set only one method (defined on a library)
included an @throws tag in its Javadoc - reporting that the given runtime exception
could be thrown in some conditions. These methods were responsible for 
267 exception stack traces mined in this study.

This result is in line with the results of two other studies~\cite{sacramento2006unchecked,kechagia2014}. 
Sacramento et al~\cite{sacramento2006unchecked} observed that the
runtime exceptions in .NET programs are most often not documented. 
Kechagia and Spinellis~\cite{kechagia2014} identified a set of methods 
on the Android API which do not document its runtime exceptions. One limitation 
of the latter work is that it did not filter out exceptions that,  
although runtime, should not be documented because they were implicitly signaled by the 
JVM due to resource restrictions or violations on Java semantic constraints.
When explicitly signaling a runtime exception and not documenting it, 
the developer imposes a threat to system robustness, especially
when such exceptions are thrown by third party code (e.g., libraries or framework utility code)
 invoked inside the application. In such cases the developer usually does not have access to 
the source code. Hence in the absence of the exception documentation it is very difficult or even impossible
 for the client to design the application to deal with ``unforeseen'' runtime exceptions. As a consequence, the
 undocumented runtime exception may remain uncaught and lead to system crashes.

\finding{%
Only a small fraction (4\%, 267 stack traces) of run time exceptions are programmatically
thrown. Almost none (0.4\%, just one) of these were documented.}

@@

 %\begin{tabular}{|p{7.9cm} p{0.1cm}|}
\begin{tabular}{lrrrr}
    \hline
 \bfseries{Origin}   &  \bfseries{stacks}  & \bfseries{signaler methods} &  \bfseries{throws clause}  &  \bfseries{@throws}  \\ 
\hline					
Libraries	& 	205	 & 	44	& 0	& 1	\\
Android  	&	62 &	35	&0 &  0	\\					
\hline					
\bfseries{All}	&	267 &	79 &  0  & 1\\
\hline					
  \end{tabular}
\caption{Absence of exception interfaces on methods.}
\label{tab:runtimeinterface}
\end{table}
